Formalization of Abstract Rewriting

### 2025-02-23 Relations/Paper/Conclusion.tex
added item
Removing missing references

Recap the main takeaways from each piece.
\begin{itemize}
  \item ARS results are effective for finitely branching and decidable relations
  --- covers first-order TRSs and the lambda calculus.
  \item Identification of ``sufficiency frontier'' for completeness (termination and confluence).
  It also collapses many well-foundedness properties.
  \item Also discuss open problems, unprovable implications, etc.
  Maybe mention the Kleene tree idea, or at least include the reference to the Bezem--Coquand paper.
  \item Throughout this effort, we found that the classical theorems often assume
  more than is necessary, and these distinctions becomes especially pronounced
  in the context of constructive logic.
  \item \emph{Formalizing classical ARS theory in the Curry--Howard framework
  lead to improvements of classical ARS results.}
\end{itemize}

>>>>

Recap the main takeaways from each piece.
\begin{itemize}
  \item ARS results are effective for finitely branching and decidable relations
  --- covers first-order TRSs and the lambda calculus.
  \item Identification of ``sufficiency frontier'' for completeness (termination and confluence).
  It also collapses many well-foundedness properties.
  \item Also discuss open problems, unprovable implications, etc.
  Maybe mention the Kleene tree idea, or at least include the reference to the Bezem--Coquand paper.
  \item Throughout this effort, we found that the classical theorems often assume
  more than is necessary, and these distinctions becomes especially pronounced
  in the context of constructive logic.
  \item \emph{Formalizing classical ARS theory in the Curry--Howard framework
  lead to improvements of classical ARS results.} In the process, new properties were identified,
   offering a clearer structural understanding of how different properties interact
    to establish completeness. These insights contribute to a more refined and
     expressive formulation of ARS theory.
  \item We believe the fact that $\isWFseq$ does not imply $\isWFacc$
  can be established by the methods of Bezem et al.

\end{itemize}

### 2025-02-23 Relations/Paper/FinalTODO.tex
Removing missing references

\begin{enumerate}
    \item Wording question. I'm trying to avoid stating "reduction" when talking about relations. However, can I still refer to a reduct and make sense? See for example the defintion of WCR.
    \item We have shown SN implies SMseq, can we show SN implies SM?
    \item Need to fill in ACM subject classifications
    \item Need to fill in keywords and phrases
    \item Need to check what other mandatory information is required in the file 
    \item Need to fill in Digital object identifier 
    \item Need to sort out a repository for the code and add it to the supplementary material section.  
    \item Add contents sections?
    \item Add bibliography/references.
    \item FINAL Remove this TODO section from the report 
\end{enumerate}

>>>>

\begin{enumerate}
    \item Need to fill in ACM subject classifications
    \item Establish [TeReSe] as how we refer to the book.
    \item Need to fill in keywords and phrases
    \item Need to check what other mandatory information is required in the file 
    \item Add bibliography/references.
    \item FINAL Remove this TODO section from the report 
\end{enumerate}

### 2025-02-23 Relations/Paper/Introduction.tex
Removing missing references

We present a formalization of the basic Abstract Rewriting Systems (ARS) theory in the Agda proof assistant.
This work is part of a larger effort to develop a library of formalized programming language theory (PL)
at Appalachian State University.
Since Term Rewriting Systems (TRS) play a foundational role in programming languages,
while ARS encompass the basic facts about rewriting relations in general,
establishing these facts is part of the necessary infrastructure required to
pursue the broader project.  The present contribution can therefore be seen as
``bootstrapping'' Appstate's new formalized PL repository.

@@

Since Strong Normalization is a central concept in ARS theory,
special attention in our work was given to various incaranations of the concept of well-foundedness in the constructive context.
(Strong normalization is well-foundedness of the converse relation.)
The standard constructive definition does not even allow one to show that SN implies WN, see Example \ref{ex:undec} \apcomment{Refer to DecidableCounterexamples.agda?}.
  We therefore looked at a number of variations of this notion, their classical counterparts all being logically equivalent.  This gave rise to a rich set of concepts, see Figure \ref{f:wf}.

>>>>

We present a formalization of the basic Abstract Rewriting Systems (ARS) theory as presented in Term Rewriting Systems [TeReSe]
 \cite{Terese}, using the Agda proof assistant.
This work is part of a larger effort to develop a library of formalized programming language theory (PL)
at Appalachian State University.
Since Term Rewriting Systems (TRS) play a foundational role in programming languages,
while ARS encompass the basic facts about rewriting relations in general,
establishing these facts is part of the necessary infrastructure required to
pursue the broader project.  The present contribution can therefore be seen as
``bootstrapping'' Appstate's new formalized PL repository.

@@

Since Strong Normalization is a central concept in ARS theory,
special attention in our work was given to various incaranations of the concept of well-foundedness in the constructive context.
(Strong normalization is well-foundedness of the converse relation.)
The standard constructive definition does not even allow one to show that SN implies WN, as we discuss in Section \ref{sec:Well-foundedness} and formalize in \texttt{WFCounters.agda}.
  We therefore looked at a number of variations of this notion, their classical counterparts all being logically equivalent. %This gave rise to a rich set of concepts, see Figure \ref{f:wf}.

### 2025-02-23 Relations/Paper/Main.tex
Removing missing references

\ccsdesc[100]{\textcolor{red}{Replace ccsdesc macro with valid one}} %TODO mandatory: Please choose ACM 2012 classifications from https://dl.acm.org/ccs/ccs_flat.cfm

\keywords{Dummy keyword} %TODO mandatory; please add comma-separated list of keywords

\supplement{https://github.com/DrPolonsky/LAM/tree/main/Relations}% TODO: set a repo just for this submissionoptional, e.g. related research data, source code, ... hosted on a repository like zenodo, figshare, GitHub, ...

@@

>>>>

\ccsdesc[300]{Theory of computation~Type theory} %TODO mandatory: Please choose ACM 2012 classifications from https://dl.acm.org/ccs/ccs_flat.cfm

\keywords{Abstract rewriting, Formalization, Confluence, Strong normalization, Newman's Lemma} %TODO mandatory; please add comma-separated list of keywords

\supplement{https://github.com/DrPolonsky/LAM//tree/main/ARS}% TODO: set a repo just for this submissionoptional, e.g. related research data, source code, ... hosted on a repository like zenodo, figshare, GitHub, ...

@@

\begin{thebibliography}{10}
    \bibitem{Terese}
    Marc Bezem, Jan Willem Klop, and Roel de Vrijer.
    \newblock {Term Rewriting Systems}.
    \newblock {Cambridge University Press, 2003}.

    \bibitem{HoTT}
    Peter Aczel, Benedikt Ahrens, Thorsten Altenkirch, Steve Awodey,  
    Bruno Barras, Andrej Bauer, Yves Bertot, Marc Bezem, Thierry Coquand,  
    Eric Finster, et al.  
    \newblock{Homotopy Type Theory: Univalent Foundations of Mathematics}.  
    \newblock{The Univalent Foundations Program, Institute for Advanced Study, 2013}.

    \bibitem{QIIT}
    Thorsten Altenkirch, Paolo Capriotti, Gabe Dijkstra, Nicolai Kraus,  
    and Fredrik Nordvall Forsberg.
    \newblock{Quotient inductive-inductive types}.  
    \newblock{Proceedings of the International Conference on Foundations of Software Science and Computation Structures},  
    \newblock{pages 293--310. Springer International Publishing, Cham, 2018}.

    \bibitem{CTT}
    Thierry Coquand, Simon Huber, and Anders Mörtberg.
    \newblock{On higher inductive types in cubical type theory}.  
    \newblock{Proceedings of the 33rd Annual ACM/IEEE Symposium on Logic in Computer Science},  
    \newblock{pages 255--264, 2018}.

    \bibitem{mimram}
    Mimram, Samuel
    \newblock{Program= Proof}.
    \newblock{2020}.

    \bibitem{Berardi}
    Marc Bezem, Thierry Coquand, Keiko Nakata, and Erik Parmann.
    \newblock{Realizability at Work: Separating Two Constructive Notions of Finiteness}.
    \newblock {In 22nd International Conference on Types for Proofs and Programs (TYPES 2016)}.
    \newblock{Leibniz International Proceedings in Informatics (LIPIcs), Volume 97, pp. 6:1-6:23, Schloss Dagstuhl – Leibniz-Zentrum für Informatik (2018)}.
    \newblock{https://doi.org/10.4230/LIPIcs.TYPES.2016.6}.

    \bibitem{Bezem}
    Stefano Berardi 
    \newblock{Classical and Intuitionistic Arithmetic with Higher Order Comprehension Coincide on Inductive Well-Foundedness}.
    \newblock{In 24th EACSL Annual Conference on Computer Science Logic (CSL 2015) }.
    \newblock{Leibniz International Proceedings in Informatics (LIPIcs), Volume 41, pp. 343-358}.
     \newblock{Schloss Dagstuhl – Leibniz-Zentrum für Informatik (2015)}.
      \newblock{https://doi.org/10.4230/LIPIcs.CSL.2015.343}.

\end{thebibliography}
